---
title: "Revisions regressions"
author: "Juraj Michalik"
date: "2024-08-13"
output: html_document
---

License:

CC-BY-4.0 by Lab of Adaptive Immunity from Institute of Molecular Genetics of the Czech Academy of Sciences

This script and all its parts are licensed under a
Creative Commons Attribution 4.0 International License.

You should have received a copy of the license along with this
work. If not, see <https://creativecommons.org/licenses/by/4.0/>.

Acknowledgements:

This project was supported by the National Institute of Virology and 
Bacteriology (Programme EXCELES, LX22NPO5103 to Ondrej Stepanek) - 
funded by the European Union - Next Generation EU.

---

This file features figures for revisions for paper of Michalik et al. Here we compute regressions on pseudobulk data generated from clusters of original data sets.

Note that file 'Figures_for_revision_tile_vln_plots.Rmd' must be run beforehand.

# Initialization

Load packages.

```{r}
library(Seurat)
library(Matrix)
library(ggplot2)
library(dplyr)
library(tidyr)
library(cowplot)
library(gridExtra)
library(ggrepel)
library(STACAS)
library(Matrix)
library(Matrix.utils)
library(scales)
library(stringr)

# negation of %in% 
`%nin%` = Negate(`%in%`)
```

```{r}
CD45RA.IDAO.col <- 'firebrick3'
CD45RO.IDAO.col <- 'darkorchid2'
CD45RA.CD45RO.IDEIS.col <- 'forestgreen'
```

Create directory where all new figures will be stored.

```{r}
dir.create('Figures_Point_by_Point', showWarnings = FALSE)
```

Load relevant data.

```{r}
data1 <- readRDS('../GSE187515/Datasets/GSE187515_integrated_data.rds')
data2 <- readRDS('../GSE212998/Datasets/GSE212998_integrated_data.rds')
```

Extract data for PTPRC assays and Antibodies for all data set and merge them into single table.

```{r}
data1.ptprc <- GetAssayData(data1, assay = 'Ptprc', slot = 'counts')
data1.atb <- GetAssayData(data1, assay = 'Antibodies', slot = 'counts')
data1.CD45er <- GetAssayData(data1, assay = 'CD45er', slot = 'counts')
data1.assays <- rbind(data1.ptprc, data1.atb, data1.CD45er) %>% t
```

```{r}
data2.ptprc <- GetAssayData(data2, assay = 'Ptprc', slot = 'counts')
data2.atb <- GetAssayData(data2, assay = 'Antibodies', slot = 'counts')
data2.CD45er <- GetAssayData(data2, assay = 'CD45er', slot = 'counts')
data2.assays <- rbind(data2.ptprc, data2.atb, data2.CD45er) %>% t
```

We generate and plot pseudobulk counts.

```{r}
clusters <- data1$seurat_clusters
data1.pseudo <- data1.assays %>% as.data.frame %>% cbind(.,clusters) %>% 
  dplyr::group_by(clusters) %>% 
  dplyr::summarise(PTPRC.RA = sum(`PTPRC-RA`), 
                   PTPRC.RO = sum(`PTPRC-RO`), 
                   CD45RA = sum(CD45RA),
                   CD45RO = sum(CD45RO),
                   erCD45RA = sum(erCD45RA),
                   erCD45RO = sum(erCD45RO))
```

Make linear regressions.

```{r}
reg1.dat1 <- lm(formula = 'PTPRC.RA ~ CD45RA', data = data1.pseudo)
reg2.dat1 <- lm(formula = 'PTPRC.RO ~ CD45RO', data = data1.pseudo)
reg3.dat1 <- lm(formula = 'PTPRC.RA ~ erCD45RA', data = data1.pseudo)
reg4.dat1 <- lm(formula = 'PTPRC.RO ~ erCD45RO', data = data1.pseudo)
```

Get correlation coeeficients.

```{r}
r.CD45RA.dat1 <- cor(data1.pseudo$CD45RA, data1.pseudo$PTPRC.RA, method = 'pearson')
r.CD45RO.dat1 <- cor(data1.pseudo$CD45RO, data1.pseudo$PTPRC.RO, method = 'pearson')
r.CD45RA.bench.dat1 <- cor(data1.pseudo$erCD45RA, data1.pseudo$PTPRC.RA, method = 'pearson')
r.CD45RO.bench.dat1 <- cor(data1.pseudo$erCD45RO, data1.pseudo$PTPRC.RO, method = 'pearson')
```

Plot pseudobulk data.

```{r}
scatterplot.lin.RA.dat1 <- ggplot(data1.pseudo %>% as.data.frame, aes(x = CD45RA, y= PTPRC.RA, label = clusters)) + 
  geom_text_repel(size = 4, max.overlaps = 100) +
  geom_point(color = CD45RA.IDAO.col) +
  theme_classic() +
  geom_abline(slope = reg1.dat1$coefficients[2], intercept = reg1.dat1$coefficients[1], col = CD45RA.IDAO.col) + 
  annotate('text',  
           col = CD45RA.IDAO.col, x = 1, y = max(data1.pseudo$PTPRC.RA), 
           label = paste0('y = ', round(reg1.dat1$coefficients[2], 2), ' + ',  round(reg1.dat1$coefficients[1], 2), '\nr = ', round(r.CD45RA.dat1,2)),
           hjust = 0) +
  ggtitle('Correspondence between Antibody and IDEIS counts \n of CD45RA isoform for GSE187515 cluster pseudobulk data') +
  xlab('CITE-seq counts') + ylab('IDEIS counts') +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'))

print(scatterplot.lin.RA.dat1)
```

```{r}
scatterplot.lin.RO.dat1 <- ggplot(data1.pseudo %>% as.data.frame, aes(x = CD45RO, y= PTPRC.RO, label = clusters)) + 
  geom_text_repel(size = 4, max.overlaps = 100) +
  geom_point(color = CD45RO.IDAO.col) +
  theme_classic() +
  geom_abline(slope = reg2.dat1$coefficients[2], intercept = reg2.dat1$coefficients[1], col = CD45RO.IDAO.col) + 
  annotate('text',  
           col = CD45RO.IDAO.col, x = 1, y = max(data1.pseudo$PTPRC.RO), 
           label = paste0('y = ', round(reg2.dat1$coefficients[2], 2), ' + ',  round(reg2.dat1$coefficients[1], 2), '\nr = ', round(r.CD45RO.dat1,2)),
           hjust = 0) +
  ggtitle('Correspondence between Antibody and IDEIS counts \n of CD45RO isoform for GSE187515 cluster pseudobulk data') +
  xlab('CITE-seq counts') + ylab('IDEIS counts') +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'))

print(scatterplot.lin.RO.dat1)
```

```{r}
scatterplot.lin.RA.bench.dat1 <- ggplot(data1.pseudo %>% as.data.frame, aes(x = erCD45RA, y= PTPRC.RA, label = clusters)) + 
  geom_text_repel(size = 4, max.overlaps = 100) +
  geom_point(color = CD45RA.IDAO.col) +
  theme_classic() +
  geom_abline(slope = reg3.dat1$coefficients[2], intercept = reg3.dat1$coefficients[1], col = CD45RA.IDAO.col) + 
  annotate('text',  
           col = CD45RA.IDAO.col, x = 1, y = max(data1.pseudo$PTPRC.RA), 
           label = paste0('y = ', round(reg3.dat1$coefficients[2], 2), ' + ',  round(reg3.dat1$coefficients[1], 2), '\nr = ', round(r.CD45RA.bench.dat1,2)),
           hjust = 0) +
  ggtitle('Correspondence between CD45er and IDEIS counts \n of CD45RA isoform for GSE187515 cluster pseudobulk data') +
  xlab('CD45er counts') + ylab('IDEIS counts') +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'))

print(scatterplot.lin.RA.bench.dat1)
```

```{r}
scatterplot.lin.RO.bench.dat1 <- ggplot(data1.pseudo %>% as.data.frame, aes(x = erCD45RO, y= PTPRC.RO, label = clusters)) + 
  geom_text_repel(size = 4, max.overlaps = 100) +
  geom_point(color = CD45RO.IDAO.col) +
  theme_classic() +
  geom_abline(slope = reg4.dat1$coefficients[2], intercept = reg4.dat1$coefficients[1], col = CD45RO.IDAO.col) + 
  annotate('text',  
           col = CD45RO.IDAO.col, x = 1, y = max(data1.pseudo$PTPRC.RO), 
           label = paste0('y = ', round(reg4.dat1$coefficients[2], 2), ' + ',  round(reg4.dat1$coefficients[1], 2), '\nr = ', round(r.CD45RO.bench.dat1,2)),
           hjust = 0) +
  ggtitle('Correspondence between CD45er and IDEIS counts \n of CD45RO isoform for GSE187515 cluster pseudobulk data') +
  xlab('CD45er counts') + ylab('IDEIS counts') +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'))

print(scatterplot.lin.RO.bench.dat1)
```
# Data sets 2

We generate and plot pseudobulk counts.

```{r}
clusters <- data2$seurat_clusters
data2.pseudo <- data2.assays %>% as.data.frame %>% cbind(.,clusters) %>% 
  dplyr::group_by(clusters) %>% 
  dplyr::summarise(PTPRC.RA = sum(`PTPRC-RA`), 
                   PTPRC.RO = sum(`PTPRC-RO`), 
                   CD45RA = sum(CD45RA),
                   CD45RO = sum(CD45RO),
                   erCD45RA = sum(erCD45RA),
                   erCD45RO = sum(erCD45RO))
```

Make linear regressions.

```{r}
reg1.dat2 <- lm(formula = 'PTPRC.RA ~ CD45RA', data = data2.pseudo)
reg2.dat2 <- lm(formula = 'PTPRC.RO ~ CD45RO', data = data2.pseudo)
reg3.dat2 <- lm(formula = 'PTPRC.RA ~ erCD45RA', data = data2.pseudo)
reg4.dat2 <- lm(formula = 'PTPRC.RO ~ erCD45RO', data = data2.pseudo)
```

get correlation coeeficients.

```{r}
r.CD45RA.dat2 <- cor(data2.pseudo$CD45RA, data2.pseudo$PTPRC.RA, method = 'pearson')
r.CD45RO.dat2 <- cor(data2.pseudo$CD45RO, data2.pseudo$PTPRC.RO, method = 'pearson')
r.CD45RA.bench.dat2 <- cor(data2.pseudo$erCD45RA, data2.pseudo$PTPRC.RA, method = 'pearson')
r.CD45RO.bench.dat2 <- cor(data2.pseudo$erCD45RO, data2.pseudo$PTPRC.RO, method = 'pearson')
```

Plot pseudobulk data.

```{r}
scatterplot.lin.RA.dat2 <- ggplot(data2.pseudo %>% as.data.frame, aes(x = CD45RA, y= PTPRC.RA, label = clusters)) + 
  geom_text_repel(size = 4, max.overlaps = 500) +
  geom_point(color = CD45RA.IDAO.col) +
  theme_classic() +
  geom_abline(slope = reg1.dat2$coefficients[2], intercept = reg1.dat2$coefficients[1], col = CD45RA.IDAO.col) + 
  annotate('text',  
           col = CD45RA.IDAO.col, x = 1, y = max(data2.pseudo$PTPRC.RA), 
           label = paste0('y = ', round(reg1.dat2$coefficients[2], 2), ' + ',  round(reg1.dat2$coefficients[1], 2), '\nr = ', round(r.CD45RA.dat2,2)),
           hjust = 0) +
  ggtitle('Correspondence between Antibody and IDEIS counts \n of CD45RA isoform for GSE212998 cluster pseudobulk data') +
  xlab('CITE-seq counts') + ylab('IDEIS counts') +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'))

print(scatterplot.lin.RA.dat2)
```

```{r}
scatterplot.lin.RO.dat2 <- ggplot(data2.pseudo %>% as.data.frame, aes(x = CD45RO, y= PTPRC.RO, label = clusters)) + 
  geom_text_repel(size = 4, max.overlaps = 500) +
  geom_point(color = CD45RO.IDAO.col) +
  theme_classic() +
  geom_abline(slope = reg2.dat2$coefficients[2], intercept = reg2.dat2$coefficients[1], col = CD45RO.IDAO.col) + 
  annotate('text',  
           col = CD45RO.IDAO.col, x = 1, y = max(data2.pseudo$PTPRC.RO), 
           label = paste0('y = ', round(reg2.dat2$coefficients[2], 2), ' + ',  round(reg2.dat2$coefficients[1], 2), '\nr = ', round(r.CD45RO.dat2,2)),
           hjust = 0) +
  ggtitle('Correspondence between Antibody and IDEIS counts \n of CD45RO isoform for GSE212998 cluster pseudobulk data') +
  xlab('CITE-seq counts') + ylab('IDEIS counts') +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'))

print(scatterplot.lin.RO.dat2)
```

```{r}
scatterplot.lin.RA.bench.dat2 <- ggplot(data2.pseudo %>% as.data.frame, aes(x = erCD45RA, y= PTPRC.RA, label = clusters)) + 
  geom_text_repel(size = 4, max.overlaps = 500) +
  geom_point(color = CD45RA.IDAO.col) +
  theme_classic() +
  geom_abline(slope = reg3.dat2$coefficients[2], intercept = reg3.dat2$coefficients[1], col = CD45RA.IDAO.col) + 
  annotate('text',  
           col = CD45RA.IDAO.col, x = 1, y = max(data2.pseudo$PTPRC.RA), 
           label = paste0('y = ', round(reg3.dat2$coefficients[2], 2), ' + ',  round(reg3.dat2$coefficients[1], 2), '\nr = ', round(r.CD45RA.bench.dat2,2)),
           hjust = 0) +
  ggtitle('Correspondence between CD45er and IDEIS counts \n of CD45RA isoform for GSE212998 cluster pseudobulk data') +
  xlab('CD45er counts') + ylab('IDEIS counts') +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'))

print(scatterplot.lin.RA.bench.dat2)
```

```{r}
scatterplot.lin.RO.bench.dat2 <- ggplot(data2.pseudo %>% as.data.frame, aes(x = erCD45RO, y= PTPRC.RO, label = clusters)) + 
  geom_text_repel(size = 4, max.overlaps = 500) +
  geom_point(color = CD45RO.IDAO.col) +
  theme_classic() +
  geom_abline(slope = reg4.dat2$coefficients[2], intercept = reg4.dat2$coefficients[1], col = CD45RO.IDAO.col) + 
  annotate('text',  
           col = CD45RO.IDAO.col, x = 1, y = max(data2.pseudo$PTPRC.RO), 
           label = paste0('y = ', round(reg4.dat2$coefficients[2], 2), ' + ',  round(reg4.dat2$coefficients[1], 2), '\nr = ', round(r.CD45RO.bench.dat2,2)),
           hjust = 0) +
  ggtitle('Correspondence between CD45er and IDEIS counts \n of CD45RO isoform for GSE212998 cluster pseudobulk data') +
  xlab('CD45er counts') + ylab('IDEIS counts') +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'))

print(scatterplot.lin.RO.bench.dat2)
```

# Data 1 - Plots for publication

Data for publication - CITE-seq and IDEIS.

```{r}
tiff('Figures_auto_revision/Supp_Figure_1/GSE187515_SupFig1_scatter_RA.tiff', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RA.dat1)
dev.off()

tiff('Figures_auto_revision/Supp_Figure_1/GSE187515_SupFig1_scatter_RO.tiff', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RO.dat1)
dev.off()
```

Save PNG for Point by Point.

```{r}
png('Figures_Point_by_Point/GSE187515_scatter_RA.png', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RA.dat1)
dev.off()

png('Figures_Point_by_Point/GSE187515_scatter_RO.png', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RO.dat1)
dev.off()
```

Create empty variants.

```{r}
scatterplot.lin.RA.blank.dat1 <- ggplot(data1.pseudo %>% as.data.frame, aes(x = CD45RA, y= PTPRC.RA)) + 
  geom_point(color = CD45RA.IDAO.col) +
  theme_classic() +
  geom_abline(slope = reg1.dat1$coefficients[2], intercept = reg1.dat1$coefficients[1], col = CD45RA.IDAO.col) + 
  ggtitle('') +
  xlab('') + ylab('') +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())
  
print(scatterplot.lin.RA.blank.dat1)
```

```{r}
scatterplot.lin.RO.blank.dat1 <- ggplot(data1.pseudo %>% as.data.frame, aes(x = CD45RO, y= PTPRC.RO)) + 
  geom_point(color = CD45RO.IDAO.col) +
  theme_classic() +
  geom_abline(slope = reg2.dat1$coefficients[2], intercept = reg2.dat1$coefficients[1], col = CD45RO.IDAO.col) + 
  ggtitle('') +
  xlab('') + ylab('') + 
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())

print(scatterplot.lin.RO.blank.dat1)
```

Save.

```{r}
tiff('Figures_auto_revision/Supp_Figure_1/GSE187515_SupFig1_scatter_RA_blank.tiff', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RA.blank.dat1)
dev.off()

tiff('Figures_auto_revision/Supp_Figure_1/GSE187515_SupFig1_scatter_RO_blank.tiff', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RO.blank.dat1)
dev.off()
```

Data for publication - CITE-seq and CD45er.

```{r}
tiff('Figures_auto_revision/Supp_Figure_1/GSE187515_SupFig1_scatter_RA_bench.tiff', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RA.bench.dat1)
dev.off()

tiff('Figures_auto_revision/Supp_Figure_1/GSE187515_SupFig1_scatter_RO_bench.tiff', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RO.bench.dat1)
dev.off()
```

Save PNG for Point by Point.

```{r}
png('Figures_Point_by_Point/GSE187515_SupFig1_scatter_RA_bench.png', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RA.dat1)
dev.off()

png('Figures_Point_by_Point/GSE187515_SupFig1_scatter_RO_bench.png', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RO.dat1)
dev.off()
```

Create empty variants.

```{r}
scatterplot.lin.RA.bench.blank.dat1 <- ggplot(data1.pseudo %>% as.data.frame, aes(x = erCD45RA, y= PTPRC.RA)) + 
  geom_point(color = CD45RA.IDAO.col) +
  theme_classic() +
  geom_abline(slope = reg3.dat1$coefficients[2], intercept = reg3.dat1$coefficients[1], col = CD45RA.IDAO.col) + 
  ggtitle('') +
  xlab('') + ylab('') +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())
  
print(scatterplot.lin.RA.bench.blank.dat1)
```

```{r}
scatterplot.lin.RO.bench.blank.dat1 <- ggplot(data1.pseudo %>% as.data.frame, aes(x = erCD45RO, y= PTPRC.RO)) + 
  geom_point(color = CD45RO.IDAO.col) +
  theme_classic() +
  geom_abline(slope = reg4.dat1$coefficients[2], intercept = reg4.dat1$coefficients[1], col = CD45RO.IDAO.col) + 
  ggtitle('') +
  xlab('') + ylab('') + 
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())

print(scatterplot.lin.RO.bench.blank.dat1)
```


Save.

```{r}
tiff('Figures_auto_revision/Supp_Figure_1/GSE187515_SupFig1_scatter_RA_blank.tiff', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RA.blank.dat1)
dev.off()

tiff('Figures_auto_revision/Supp_Figure_1/GSE187515_SupFig1_scatter_RO_blank.tiff', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RO.blank.dat1)
dev.off()

tiff('Figures_auto_revision/Supp_Figure_1/GSE187515_SupFig1_scatter_RA_blank_bench.tiff', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RA.bench.blank.dat1)
dev.off()

tiff('Figures_auto_revision/Supp_Figure_1/GSE187515_SupFig1_scatter_RO_blank_bench.tiff', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RO.bench.blank.dat1)
dev.off()
```

```{r}
png('Figures_Point_by_Point/GSE187515_SupFig1_scatter_RA_bench.png', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RA.dat1)
dev.off()

png('Figures_Point_by_Point/GSE187515_SupFig1_scatter_RO_bench.png', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RO.dat1)
dev.off()
```

# Data 2

```{r}
png('Figures_Point_by_Point/GSE212998_SupFig1_scatter_RA.png', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RA.dat2)
dev.off()

png('Figures_Point_by_Point/GSE212998_SupFig1_scatter_RO.png', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RO.dat2)
dev.off()
```

```{r}
png('Figures_Point_by_Point/GSE212998_SupFig1_scatter_RA_bench.png', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RA.bench.dat2)
dev.off()

png('Figures_Point_by_Point/GSE212998_SupFig1_scatter_RO_bench.png', width = 6, height = 6, units = 'in', res = 300)
print(scatterplot.lin.RO.bench.dat2)
dev.off()
```
