---
title: "Revisions - comparing intergation with CITE-seq and IDEIS"
author: "Juraj Michalik"
date: "2024-08-11"
output: html_document
---

License:

CC-BY-4.0 by Lab of Adaptive Immunity from Institute of Molecular Genetics of the Czech Academy of Sciences

This script and all its parts are licensed under a
Creative Commons Attribution 4.0 International License.

You should have received a copy of the license along with this
work. If not, see <https://creativecommons.org/licenses/by/4.0/>.

Acknowledgements:

This project was supported by the National Institute of Virology and 
Bacteriology (Programme EXCELES, LX22NPO5103 to Ondrej Stepanek) - 
funded by the European Union - Next Generation EU.

---

This file features figures for Point by Point for paper of Michalik et al. Here we compute UMAP visualisations using CD45RA and CD45RO as determinedby either CITE-seq or IDEIS as features.

Note that file 'Pseudobuk_R_regressons.Rmd' must be run beforehand.

# Initialization

Load packages.

```{r}
library(Seurat)
library(Matrix)
library(ggplot2)
library(dplyr)
library(tidyr)
library(cowplot)
library(gridExtra)
library(ggrepel)
library(STACAS)
library(Matrix)
library(Matrix.utils)
library(scales)
library(stringr)

# negation of %in% 
`%nin%` = Negate(`%in%`)
```

Load data.

```{r}
data2004 <- readRDS('../GSE187515/Datasets/dataset_2004_finalized_data.rds')
data2023rep1 <- readRDS('../GSE187515/Datasets/dataset_2023Rep1_finalized_data.rds')
data2023rep2 <- readRDS('~../GSE187515/Datasets/dataset_2023Rep2_finalized_data.rds')
  
HIV_list <- list(data2004, data2023rep1, data2023rep2)
```

Create directory where new RDS will be saved.

```{r}
dir.create('Datasets')
```  
  
# Passing CITE-seq and IDEIS as features

We pass data to each assay from CITE seq as new feature.

```{r}
counts.2004 <- GetAssayData(HIV_list[[1]], slot = 'counts', assay = 'RNA')
counts.2023rep1 <- GetAssayData(HIV_list[[2]], slot = 'counts' , assay = 'RNA')
counts.2023rep2 <- GetAssayData(HIV_list[[3]], slot = 'counts' , assay = 'RNA')

counts.2004.CITE <- GetAssayData(HIV_list[[1]], slot = 'counts', assay = 'Antibodies')
counts.2023rep1.CITE <- GetAssayData(HIV_list[[2]], slot = 'counts', assay = 'Antibodies')
counts.2023rep2.CITE <- GetAssayData(HIV_list[[3]], slot = 'counts', assay = 'Antibodies')

counts.2004.IDEIS <- GetAssayData(HIV_list[[1]], slot = 'counts', assay = 'Ptprc')
counts.2023rep1.IDEIS <- GetAssayData(HIV_list[[2]], slot = 'counts', assay = 'Ptprc')
counts.2023rep2.IDEIS <- GetAssayData(HIV_list[[3]], slot = 'counts', assay = 'Ptprc')

# subfilter to genes of interest

counts.2004.CITE <- counts.2004.CITE[c('CD45RA', 'CD45RO'),]
counts.2023rep1.CITE <- counts.2023rep1.CITE[c('CD45RA', 'CD45RO'),]
counts.2023rep2.CITE <- counts.2023rep2.CITE[c('CD45RA', 'CD45RO'),]

counts.2004.IDEIS <- counts.2004.IDEIS[c('PTPRC-RA', 'PTPRC-RO'),]
counts.2023rep1.IDEIS <- counts.2023rep1.IDEIS[c('PTPRC-RA', 'PTPRC-RO'),]
counts.2023rep2.IDEIS <- counts.2023rep2.IDEIS[c('PTPRC-RA', 'PTPRC-RO'),]
```

# Passing CITE-seq as features

Merge data, make objects and make list.

```{r}
counts.2004.CITE.GEX <- rbind(counts.2004, counts.2004.CITE)
counts.2023rep1.CITE.GEX <- rbind(counts.2023rep1, counts.2023rep1.CITE)
counts.2023rep2.CITE.GEX <- rbind(counts.2023rep2, counts.2023rep2.CITE)

seu.2004.CITE <- CreateSeuratObject(counts.2004.CITE.GEX)
seu.2023.rep1.CITE <- CreateSeuratObject(counts.2023rep1.CITE.GEX)
seu.2023.rep2.CITE <- CreateSeuratObject(counts.2023rep2.CITE.GEX)

seu.2004.CITE[['Ptprc']] <- CreateAssayObject(counts.2004.IDEIS)
seu.2023.rep1.CITE[['Ptprc']] <- CreateAssayObject(counts.2023rep1.IDEIS)
seu.2023.rep2.CITE[['Ptprc']] <- CreateAssayObject(counts.2023rep2.IDEIS)

HIV.CITE <- list(seu.2004.CITE, seu.2023.rep1.CITE, seu.2023.rep2.CITE)
```

Now we need to renormalize and select variable features.

```{r}
for(i in 1:length(HIV.CITE)){
  HIV.CITE[[i]] <- NormalizeData(HIV.CITE[[i]])
  HIV.CITE[[i]] <- FindVariableFeatures(HIV.CITE[[i]], selection.method = "vst",
                                        nfeatures = 1000, verbose = FALSE)
}
```

```{r}
HIV.CITE.integ <- Run.STACAS(HIV.CITE, anchor.features = c(VariableFeatures(HIV.CITE[[1]]), 'CD45RA', 'CD45RO'))
```

```{r}
HIV.CITE.integ <- ScaleData(HIV.CITE.integ, verbose = FALSE)
HIV.CITE.integ <- RunPCA(HIV.CITE.integ, npcs = 20, verbose = FALSE)
HIV.CITE.integ <- RunUMAP(HIV.CITE.integ, reduction = "pca", dims = 1:20)
  
HIV.CITE.integ <- FindNeighbors(HIV.CITE.integ, dims = 1:20)
HIV.CITE.integ <- FindClusters(HIV.CITE.integ, resolution = 0.5)
```

```{r}
DimPlot(HIV.CITE.integ, label = T) + ggtitle('HIV-CD4+ analysis with CITE-seq as features')
VlnPlot(HIV.CITE.integ, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2, pt.size = 0)
```

```{r}
png('Figures_Point_by_Point//GSE187515_UMAP_with_CITEseq.png', width = 6.5, height = 6, units = 'in', res = 200)
print(DimPlot(HIV.CITE.integ, label = T) + ggtitle('HIV-CD4+ analysis with CITE-seq as features'))
dev.off()
```

```{r}
png('Figures_Point_by_Point/GSE187515_UMAP_with_CITEseq_FeaturePlot.png', width = 11, height = 6, units = 'in', res = 200)
print(FeaturePlot(HIV.CITE.integ, features = c('CD45RA', 'CD45RO'), min.cutoff = 0))
dev.off()
```

Save.

```{r}
saveRDS(HIV.CITE.integ, 'Datasets/HIV_with_CITE.rds')
```

# Passing IDEIS as feature

We pass data to each assay from IDEIS seq as new feature.

Merge data, make objects and make list.

```{r}
counts.2004.IDEIS.GEX <- rbind(counts.2004, counts.2004.IDEIS)
counts.2023rep1.IDEIS.GEX <- rbind(counts.2023rep1, counts.2023rep1.IDEIS)
counts.2023rep2.IDEIS.GEX <- rbind(counts.2023rep2, counts.2023rep2.IDEIS)

seu.2004.IDEIS <- CreateSeuratObject(counts.2004.IDEIS.GEX)
seu.2023.rep1.IDEIS <- CreateSeuratObject(counts.2023rep1.IDEIS.GEX)
seu.2023.rep2.IDEIS <- CreateSeuratObject(counts.2023rep2.IDEIS.GEX)

seu.2004.IDEIS[['Antibodies']] <- CreateAssayObject(counts.2004.CITE)
seu.2023.rep1.IDEIS[['Antibodies']] <- CreateAssayObject(counts.2023rep1.CITE)
seu.2023.rep2.IDEIS[['Antibodies']] <- CreateAssayObject(counts.2023rep2.CITE)

HIV.IDEIS <- list(seu.2004.IDEIS, seu.2023.rep1.IDEIS, seu.2023.rep2.IDEIS)
```

Now we need to renormalize and select variable features.

```{r}
for(i in 1:length(HIV.IDEIS)){
  HIV.IDEIS[[i]] <- NormalizeData(HIV.IDEIS[[i]])
  HIV.IDEIS[[i]] <- FindVariableFeatures(HIV.IDEIS[[i]], selection.method = "vst",
                                        nfeatures = 1000, verbose = FALSE)
}
```

```{r}
HIV.IDEIS.integ <- Run.STACAS(HIV.IDEIS, anchor.features = c(VariableFeatures(HIV.IDEIS[[1]]), 'PTPRC-RA', 'PTPRC-RO'))
```

```{r}
HIV.IDEIS.integ <- ScaleData(HIV.IDEIS.integ, verbose = FALSE)
HIV.IDEIS.integ <- RunPCA(HIV.IDEIS.integ, npcs = 20, verbose = FALSE)
HIV.IDEIS.integ <- RunUMAP(HIV.IDEIS.integ, reduction = "pca", dims = 1:20)
  
HIV.IDEIS.integ <- FindNeighbors(HIV.IDEIS.integ, dims = 1:20)
HIV.IDEIS.integ <- FindClusters(HIV.IDEIS.integ, resolution = 0.5)
```

```{r}
DimPlot(HIV.IDEIS.integ, label = T) + ggtitle('HIV-CD4+ analysis with IDEIS as features')
VlnPlot(HIV.IDEIS.integ, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2, pt.size = 0)
```

```{r}
png('Figures_Point_by_Point/GSE187515_UMAP_with_IDEIS.png', width = 6.5, height = 6, units = 'in', res = 200)
print(DimPlot(HIV.IDEIS.integ, label = T) + ggtitle('HIV-CD4+ analysis with IDEIS as features'))
dev.off()
```

```{r}
png('Figures_Point_by_Point/GSE187515_UMAP_with_IDEIS_FeaturePlot.png', width = 11, height = 6, units = 'in', res = 200)
print(FeaturePlot(HIV.IDEIS.integ, features = c('PTPRC-RA', 'PTPRC-RO'), min.cutoff = 0))
dev.off()
```
Save.

```{r}
saveRDS(HIV.IDEIS.integ, 'Datasets/HIV_with_IDEIS.rds')
```
