---
title: "GSE155006"
author: "Juraj Michalik"
date: "2023-07-25"
output: html_document
---

License:

CC-BY-4.0 by Lab of Adaptive Immunity from Institute of Molecular Genetics of the Czech Academy of Sciences

This script and all its parts are licensed under a
Creative Commons Attribution 4.0 International License.

You should have received a copy of the license along with this
work. If not, see <https://creativecommons.org/licenses/by/4.0/>.

Acknowledgements:

This project was supported by the National Institute of Virology and 
Bacteriology (Programme EXCELES, LX22NPO5103 to Ondrej Stepanek) - 
funded by the European Union - Next Generation EU.

---

Analysis of Cd45/Ptprc isoforms from CD45+ cells from various murine organs. These data lack FB so we need to use standard markers to identify cell types and see the match with Cd45 isoform expression.
Paper: https://www.sciencedirect.com/science/article/pii/S1074761320304921?via%3Dihub

# Initialization

## Creating objects

Load libraries.

```{r setup, include=FALSE}
library(Seurat)
library(future)
library(ggplot2)
library(dplyr)
library(tidyr)
library(cowplot)
library(gridExtra)
library(dbscan)
library(scatterpie)
library(reshape2)
library(ggrepel)
library(scds)
library(SingleCellExperiment)
library(STACAS)
library(SingleR)
library(celldex)

# negation of %in% 
`%nin%` = Negate(`%in%`)
```

Create required directories.

```{r}
dir.create('Datasets', showWarnings = FALSE)
dir.create('Whitelists', showWarnings = FALSE)
```

Load all data (8 samples).

```{r}
count.mouse.liver.aged <- Read10X('Count_matrices/GSM4321524_Liver_Aged/')
count.mouse.liver.young <- Read10X('Count_matrices/GSM4321525_Liver_Young/')
count.mouse.lungs.aged <- Read10X('Count_matrices/GSM4321526_Lungs_Aged/')
count.mouse.lungs.young <- Read10X('Count_matrices/GSM4321527_Lungs_Young/')
count.mouse.peritoneal.cells.aged <- Read10X('Count_matrices/GSM4321528_Peritoneal_cells_Aged/')
count.mouse.peritoneal.cells.young <- Read10X('Count_matrices/GSM4321529_Peritoneal_cells_Young/')
count.mouse.spleen.aged <- Read10X('Count_matrices/GSM4321530_Spleen_Aged/')
count.mouse.spleen.young <- Read10X('Count_matrices/GSM4321531_Spleen_Young/')
```

```{r}
mm.liver.aged <- CreateSeuratObject(count.mouse.liver.aged, project = 'Aged Liver')
mm.liver.young <- CreateSeuratObject(count.mouse.liver.young, project = 'Young Liver')
mm.lungs.aged <- CreateSeuratObject(count.mouse.lungs.aged, project = 'Aged Lungs')
mm.lungs.young <- CreateSeuratObject(count.mouse.lungs.young, project = 'Young Lungs')
mm.peritoneal.cells.aged <- CreateSeuratObject(count.mouse.peritoneal.cells.aged, project = 'Aged Peritoneal Cells')
mm.peritoneal.cells.young <- CreateSeuratObject(count.mouse.peritoneal.cells.young, project = 'Young Peritoneal Cells')
mm.spleen.aged <- CreateSeuratObject(count.mouse.spleen.aged, project = 'Aged Spleen')
mm.spleen.young <- CreateSeuratObject(count.mouse.spleen.young, project = 'Young Spleen')
```

Add percentage of MT and RP genes, then remove said genes.

```{r, warning=F, message=F}
mm.mog.list <- list(mm.liver.aged, mm.liver.young, mm.lungs.aged, mm.lungs.young,
                    mm.peritoneal.cells.aged, mm.peritoneal.cells.young, mm.spleen.aged, mm.spleen.young)
```

```{r}
for(i in 1:length(mm.mog.list)){
  mm.mog.list[[i]][["percent.mt"]] <- PercentageFeatureSet(object = mm.mog.list[[i]], pattern = "^mt-")
  mm.mog.list[[i]][["percent.rp"]] <- PercentageFeatureSet(object = mm.mog.list[[i]], pattern = "^Rp[s|l]")  
}
```

Now filter out those genes as well as genes related to V(D)J.

```{r}
for(i in 1:length(mm.mog.list)){
  mm.mog.list[[i]] <- mm.mog.list[[i]][!(grepl('^Tr[a|b][v|d|j]', rownames(mm.mog.list[[i]])) |
                                         grepl('^mt-', rownames(mm.mog.list[[i]])) |
                                         grepl('^Rp[s|l]', rownames(mm.mog.list[[i]]))),]
}
```

```{r}
mog.names <- c('Liver_Aged', 'Liver_Young', 'Lungs_Aged', 'Lungs_Young',
               'Peritoneal_Cells_Aged', 'Peritoneal_Cells_Young', 'Spleen_Aged', 'Spleen_Young')
```

Write whitelists.

```{r}
for(i in 1:length(mm.mog.list)){
  write.table(gsub('-1','', colnames(mm.mog.list[[i]])), paste0('Whitelists/Whitelist_', mog.names[i], '.csv'), quote = F, row.names = F, col.names = F)
}
```

Save initial data sets.

```{r}
for(i in 1:length(mm.mog.list)){
  saveRDS(mm.mog.list[[i]], paste0('Datasets/', mog.names[i], '_initial_data.rds'))
}
```

Now run IDEIS and follow with corresponding analysis file.
