---
title: "GSE155006"
author: "Juraj Michalik"
date: "2023-07-25"
output: html_document
---

License:

CC-BY-4.0 by Lab of Adaptive Immunity from Institute of Molecular Genetics of the Czech Academy of Sciences

This script and all its parts are licensed under a
Creative Commons Attribution 4.0 International License.

You should have received a copy of the license along with this
work. If not, see <https://creativecommons.org/licenses/by/4.0/>.

Acknowledgements:

This project was supported by the National Institute of Virology and 
Bacteriology (Programme EXCELES, LX22NPO5103 to Ondrej Stepanek) - 
funded by the European Union - Next Generation EU.

---

Analysis of Cd45/Ptprc isoforms from CD45+ cells from various murine organs. These data lack FB so we need to use standard markers to identify cell types and see the match with Cd45 isoform expression.
Paper: https://www.sciencedirect.com/science/article/pii/S1074761320304921?via%3Dihub

Run this script once GSE155006_preparation.Rmd and IDEIS were run.

# Initialization

## Creating objects

Load libraries.

```{r setup, include=FALSE}
library(Seurat)
library(future)
library(ggplot2)
library(dplyr)
library(tidyr)
library(cowplot)
library(gridExtra)
library(dbscan)
library(scatterpie)
library(reshape2)
library(ggrepel)
library(scds)
library(SingleCellExperiment)
library(STACAS)
library(SingleR)
library(celldex)

# negation of %in% 
`%nin%` = Negate(`%in%`)
```

```{r}
CD45RA.col <- 'firebrick3'
CD45RB.col <- 'green4'
CD45RC.col <- 'gray20'
CD45RO.col <- 'darkorchid2'

CD45RB.CD45RO.IDEIS.col <- 'forestgreen'
```

Create required directories.

```{r}
dir.create('Figure_5', showWarnings = FALSE)
```

Load data initialized beforehand (8 samples).

```{r}
mm.mog.list <- list()
names.files <- c('Liver_Aged', 'Liver_Young', 'Lungs_Aged', 'Lungs_Young',
                 'Peritoneal_Cells_Aged', 'Peritoneal_Cells_Young', 'Spleen_Aged', 'Spleen_Young')

for(i in 1:length(names.files)){
  mm.mog.list[[i]] <- readRDS(paste0('Datasets/', names.files[i], '_initial_data.rds'))
}
names(mm.mog.list) <- names.files
```

```{r}
for(i in 1:length(mm.mog.list)){
  mm.mog.ptprc <- readRDS(paste0('Ptprc/Ptprc_', names(mm.mog.list)[i], '.rds'))
  mm.mog.list[[i]][['Ptprc']] <- CreateAssayObject(mm.mog.ptprc)
}
```

```{r}
for(i in 1:length(mm.mog.list)){
  print(VlnPlot(mm.mog.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rp"), ncol = 4, pt.size = 0))
}
```

```{r warning=F, message=F}
for(i in 1:length(mm.mog.list)){
  mm.mog.list[[i]] <- NormalizeData(object = mm.mog.list[[i]])
  mm.mog.list[[i]] <- FindVariableFeatures(mm.mog.list[[i]], selection.method = "vst",
                                        nfeatures = 1000, verbose = FALSE)
    
  mm.mog.list[[i]] <- ScaleData(mm.mog.list[[i]], verbose = FALSE)
  mm.mog.list[[i]] <- RunPCA(mm.mog.list[[i]], npcs = 20, verbose = FALSE)
  mm.mog.list[[i]] <- RunUMAP(mm.mog.list[[i]], reduction = "pca", dims = 1:20)
    
  mm.mog.list[[i]] <- FindNeighbors(mm.mog.list[[i]], dims = 1:20)
  mm.mog.list[[i]] <- FindClusters(mm.mog.list[[i]], resolution = 0.5)
}
```


```{r}
titles <- c('MM Aged Liver', 'MM Young Liver', 'MM Aged Lungs', 'MM Young Lungs', 'MM Aged Peritoneal cells', 'MM Young Peritoneal Cells', 'MM Aged Spleen', 'MM Young Spleen')
for(i in 1:length(mm.mog.list)){
  DimPlot(mm.mog.list[[i]], label = T) + ggtitle(titles[[i]])
  VlnPlot(mm.mog.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rp"), ncol = 4, pt.size = 0)
}
```

```{r, warning = F, fig.width = 10}
for(i in 1:length(mm.mog.list)){
  print(FeaturePlot(mm.mog.list[[i]], c('Ptprc-RA', 'Ptprc-RB', 'Ptprc-RC', 'Ptprc-RO')))
}
```

# Cleaning data

We now prepare and clean data. First we perform DEA to see some genes for each data set and cluster.

```{r}
DEA.data <- list()

for(i in 1:length(mm.mog.list)){
  DEA.set <- FindAllMarkers(mm.mog.list[[i]], assay = 'RNA')
  DEA.set <- DEA.set %>% dplyr::group_by(cluster) %>% dplyr::filter(row_number() <= 20)
  DEA.data[[i]] <- DEA.set
}
```


```{r}
print(DEA.data)
```

We show some of markers for each case.

```{r, fig.width = 10}
for(i in 1:length(mm.mog.list)){
  print(FeaturePlot(mm.mog.list[[i]], c('Cd8a', 'Cd8b1', 'Cd4', 'Ptprc','Cd79a', 'Mki67'), ncol = 3))
}
```

Now we remove following:

* Cluster 10 from Aged Lives samples (lacks CD45);
* Cluster 1 from Young Liver samples (lacks CD45), and cluster 5 (dead);
* Cluster 9 from Aged Lung samples (dead cells);
* Cluster 11 (dead) and 14, 15 from Young Lung samples (lacks CD45);
* Cluster 4 from Aged Peritoneal Cell samples (dead);
* Cluster 8 from Young Peritoneal Cell samples (dead);
* Cluster 1 from Aged Spleen samples (dead);
* Clusters 2 and 7 from Young Spleen samples (lacks CD45).
 
We then follow-up with re-mapping and clustering. All cells with more than 10% of gene mapping to MT will be also removed.

```{r warning=F, message=F}
mm.mog.list[[1]] <- subset(mm.mog.list[[1]], seurat_clusters %nin% c(10) & percent.mt <= 10)
mm.mog.list[[2]] <- subset(mm.mog.list[[2]], seurat_clusters %nin% c(1,5) & percent.mt <= 10)
mm.mog.list[[3]] <- subset(mm.mog.list[[3]], seurat_clusters %nin% c(9) & percent.mt <= 10)
mm.mog.list[[4]] <- subset(mm.mog.list[[4]], seurat_clusters %nin% c(11, 14, 15) & percent.mt <= 10)
mm.mog.list[[5]] <- subset(mm.mog.list[[5]], seurat_clusters %nin% c(4) & percent.mt <= 10)
mm.mog.list[[6]] <- subset(mm.mog.list[[6]], seurat_clusters %nin% c(8) & percent.mt <= 10)
mm.mog.list[[7]] <- subset(mm.mog.list[[7]], seurat_clusters %nin% c(1) & percent.mt <= 10)
mm.mog.list[[8]] <- subset(mm.mog.list[[8]], seurat_clusters %nin% c(2, 7) & percent.mt <= 10)
```

```{r}
for(i in 1:length(mm.mog.list)){
  mm.mog.list[[i]] <- NormalizeData(object = mm.mog.list[[i]])
  mm.mog.list[[i]] <- FindVariableFeatures(mm.mog.list[[i]], selection.method = "vst",
                                           nfeatures = 1000, verbose = FALSE)
    
  mm.mog.list[[i]] <- ScaleData(mm.mog.list[[i]], verbose = FALSE)
  mm.mog.list[[i]] <- RunPCA(mm.mog.list[[i]], npcs = 20, verbose = FALSE)
  mm.mog.list[[i]] <- RunUMAP(mm.mog.list[[i]], reduction = "pca", dims = 1:20)
    
  mm.mog.list[[i]] <- FindNeighbors(mm.mog.list[[i]], dims = 1:20)
  mm.mog.list[[i]] <- FindClusters(mm.mog.list[[i]], resolution = 0.5)
}
```

Normalize assays with Ptprc.

```{r}
for(i in 1:length(mm.mog.list)){
  mm.mog.list[[i]] <- NormalizeData(mm.mog.list[[i]], normalization.method = "CLR", margin = 2, assay = "Ptprc")
}
```

```{r}
fig.titles <- c('MM Aged Liver', 'MM Young Liver', 'MM Aged Lungs', 'MM Young Lungs',
                'MM Aged Peritoneal Cells', 'MM Young Peritoneal Cells', 'MM Aged Spleen', 'MM Young Spleen')
for(i in 1:length(mm.mog.list)){
  DimPlot(mm.mog.list[[i]], label = T) + ggtitle(fig.titles[i])
  VlnPlot(mm.mog.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rp"), ncol = 4, pt.size = 0)
}
```

```{r, warning = F, fig.width = 10}
for(i in 1:length(mm.mog.list)){
  FeaturePlot(mm.mog.list[[i]], c('Ptprc-RA', 'Ptprc-RB', 'Ptprc-RC', 'Ptprc-RO'))
}
```

Save all sample data sets.

```{r}
for(i in 1:length(mm.mog.list)){
  saveRDS(mm.mog.list[[i]], paste0('Datasets/', names(mm.mog.list)[i], '_finalized_data.rds'))
}
```

# Integration

We now integrate the data.

```{r}
mm.mog.integ.data <- Run.STACAS(mm.mog.list, anchor.features = 1000)
```

```{r}
mm.mog.integ.data <- ScaleData(mm.mog.integ.data, verbose = FALSE)
mm.mog.integ.data <- RunPCA(mm.mog.integ.data, npcs = 20, verbose = FALSE)
mm.mog.integ.data <- RunUMAP(mm.mog.integ.data, reduction = "pca", dims = 1:20)
  
mm.mog.integ.data <- FindNeighbors(mm.mog.integ.data, dims = 1:20)
mm.mog.integ.data <- FindClusters(mm.mog.integ.data, resolution = 0.8)
```

```{r}
DimPlot(mm.mog.integ.data, label = T) + ggtitle('CD45+ T cells in Young and Aged mice data integration')
VlnPlot(mm.mog.integ.data, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rp"), ncol = 4, pt.size = 0)
DimPlot(mm.mog.integ.data, group.by = 'orig.ident') + ggtitle('CD45+ T cells in Young and Aged mice data integration')
```

```{r, warning = F, fig.width = 10}
FeaturePlot(mm.mog.integ.data, c('Ptprc-RA', 'Ptprc-RB', 'Ptprc-RC', 'Ptprc-RO'))
```

Save integrated data sets.

```{r}
saveRDS(mm.mog.integ.data, 'Datasets/GSE155006_integrated_data.rds')
```

```{r}
mm.mog.DEA.integrated <- FindAllMarkers(mm.mog.integ.data, assay = 'RNA', logfc.threshold = 1)
```

```{r}
print(mm.mog.DEA.integrated %>% dplyr::group_by(cluster) %>% dplyr::filter(row_number() <= 20))
```

Print some interesting genes.

```{r, fig.width = 14}
FeaturePlot(mm.mog.integ.data, c('Ptprc', 'Cd79a', 'Cd79b', 'Cd3d', 'Cd3e', 'Cd4', 'Cd8a', 'Cd8b1', 'Ms4a1', 'C1qc'), ncol = 4, min.cutoff = 0)
VlnPlot(mm.mog.integ.data, c('Ptprc', 'Cd8a', 'Cd8b1', 'Cd4', 'Ms4a1'), ncol = 3)
```

We notice following:

* Cluster 25 lacks Ptprc and shows presence of genes such as Cyp11a1 typical for adrenals. Most likely these are not immune cells, so we reject them.
* Cluster 2 lacks any interesting markers beyond infection so we'll remove it as well.
* Clusters 3 and 18 have significant lack of Cd45/Ptprc and presence of macrophage markers, so they'll get eliminated as well.
* Cluster 19 contains mainly proliferating cells and lacks Ptprc.
* Cluster 15 contains markers for both B-cells and T-cells, so we remove this cluster as well.
* Clusters 20, 23 and 24 are small and also relatively low on Ptprc count, plus they're hard to identify.
* Cluster 22 is small and hard to identify, so it gets removed as well.

We remove all cells then perform splitting, re-normalization and reintegration.

```{r}
mm.mog.integ.data.all <- mm.mog.integ.data
```

```{r}
mm.mog.integ.data <- subset(mm.mog.integ.data, seurat_clusters %nin% c(2, 3, 15, 18, 19, 20, 22, 23, 24, 25))
```

```{r}
mm.mog.list.split <- SplitObject(mm.mog.integ.data, split.by = 'orig.ident')
for(i in 1:length(mm.mog.list.split)){
  DefaultAssay(mm.mog.list.split[[i]]) <- 'RNA'
  mm.mog.list.split[[i]] <- NormalizeData(object = mm.mog.list.split[[i]])

  mm.mog.list.split[[i]] <- NormalizeData(mm.mog.list.split[[i]], normalization.method = "CLR", margin = 2, assay = "Ptprc")
}
```

Now reintegrate.

```{r}
mm.mog.integ.data <- Run.STACAS(mm.mog.list.split, anchor.features = 700)
```

```{r}
mm.mog.integ.data <- ScaleData(mm.mog.integ.data, verbose = FALSE)
mm.mog.integ.data <- RunPCA(mm.mog.integ.data, npcs = 20, verbose = FALSE)
mm.mog.integ.data <- RunUMAP(mm.mog.integ.data, reduction = "pca", dims = 1:20)
  
mm.mog.integ.data <- FindNeighbors(mm.mog.integ.data, dims = 1:20)
mm.mog.integ.data <- FindClusters(mm.mog.integ.data, resolution = 0.3)
```

```{r}
DimPlot(mm.mog.integ.data, label = T) + ggtitle('CD45+ T cells in Young and Aged mice data integration')
VlnPlot(mm.mog.integ.data, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rp"), ncol = 4, pt.size = 0)
DimPlot(mm.mog.integ.data, group.by = 'orig.ident') + ggtitle('CD45+ T cells in Young and Aged mice data integration')
```

```{r, warning = F, fig.width = 14}
FeaturePlot(mm.mog.integ.data, c('Ptprc-RA', 'Ptprc-RB', 'Ptprc-RC', 'Ptprc-RO', 'Ptprc'), ncol = 3)
```

```{r}
mm.mog.DEA.integrated.filt <- FindAllMarkers(mm.mog.integ.data, assay = 'RNA', logfc.threshold = 1)
```

```{r}
print(mm.mog.DEA.integrated.filt %>% dplyr::group_by(cluster) %>% dplyr::filter(row_number() <= 20))
```

Save updated integrated data set.

```{r}
saveRDS(mm.mog.integ.data, 'Datasets/GSE155006_integrated_data.rds')
```

# Generating figures

Generate final figures. For each figure we generate Desired final plot and equivalent without text. 

DimPlot:

```{r}
dim.plot.1 <- DimPlot(mm.mog.integ.data, label = T) + 
  ggtitle('PBMC (GSE155006) - Clustering') + 
  scale_x_continuous(breaks=c(-15, 0, 10)) +
  scale_y_continuous(breaks=c(-15, 0, 5)) +
  guides(color = guide_legend(byrow = TRUE,
                              override.aes = list(size=6))) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.spacing.x = unit(1, 'cm'),
        legend.spacing.y = unit(0.3, 'cm'))
print(dim.plot.1) 
```

```{r}
tiff('Figure_5/GSE155006_DimPlot.tiff', width = 8.5, height = 6, units = 'in', res = 200)
print(dim.plot.1)
dev.off()
```

```{r}
dim.plot.1.blank <- DimPlot(mm.mog.integ.data) + 
  ggtitle('') + xlab('') + ylab('') +
  scale_x_continuous(breaks=c(-15, 0, 10)) +
  scale_y_continuous(breaks=c(-15, 0, 5)) +
  guides(color = guide_legend(byrow = TRUE,
                              override.aes = list(size=6))) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        legend.title=element_blank(),
        legend.text=element_blank(),
        legend.spacing.y = unit(0.3, 'cm'),
        legend.spacing.x = unit(1, 'cm'))
print(dim.plot.1.blank) 
```

```{r}
tiff('Figure_5/GSE155006_DimPlot_blank.tiff', width = 8, height = 6, units = 'in', res = 200)
print(dim.plot.1.blank)
dev.off()
```

Feature plots: PTPRC-RA/RO (Features)

```{r, message= F, warning = F}
feature.to.list <- c('Ptprc-RA', 'Ptprc-RB', 'Ptprc-RC', 'Ptprc-RO')
colors <- c(CD45RA.col, CD45RB.col, CD45RC.col, CD45RO.col)
name.tags <- c('CD45RA', 'CD45RB', 'CD45RC', 'CD45RO')
lims <- c(1,1,0.5,1)

for(i in 1:length(feature.to.list)){
  fb.plot <- FeaturePlot(mm.mog.integ.data, feature.to.list[i], min.cutoff = 0, pt.size = 0.75) + 
    ggtitle(paste0('PBMC (GSE155006) - ', name.tags[i], ' -IdAO')) + 
    scale_colour_gradient(breaks = c(0,lims[i]), low = 'grey85', high = colors[i]) +
    scale_x_continuous(breaks=c(-15, 0, 10)) +
    scale_y_continuous(breaks=c(-15, 0, 5)) +
    theme(plot.title = element_text(hjust = 0.5))
  print(fb.plot) 
  
  tiff(paste0('Figure_5/GSE155006_FeaturePlot_', name.tags[i], '_Software.tiff'), width = 7, height = 6, units = 'in', res = 200)
  print(fb.plot)
  dev.off()
  
  fb.plot.blank <- FeaturePlot(mm.mog.integ.data, feature.to.list[i], min.cutoff = 0, pt.size = 0.75) + 
    ggtitle('') + xlab('') + ylab('') +
    scale_colour_gradient(breaks = c(0,lims[i]), low = 'grey85', high = colors[i]) +
    scale_x_continuous(breaks=c(-15, 0, 10)) +
    scale_y_continuous(breaks=c(-15, 0, 5)) +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x=element_blank(),
          axis.text.y=element_blank(),
          legend.title=element_blank(),
          legend.text=element_blank())
  print(fb.plot.blank) 
  
  tiff(paste0('Figure_5/GSE155006_FeaturePlot_', name.tags[i], '_Software_blank.tiff'), width = 6.5, height = 6, units = 'in', res = 200)
  print(fb.plot.blank)
  dev.off()
}
```

Plot desired features.

```{r, message=F, warning=F}
features.to.plot = c('Cd28', 'Cd79a', 'Cd3d', 'Cd4', 'Cd8a', 'Cd14', 'Klrk1', 'Sell', 'Ccr7', 'Ccl5', 'Cd44', 'Tcrg-C1', 'Trgv2', 'H2-Ab1')
breaks <- c(4,5,4,3,4,2,4,4,4,7,4,4,4,5)
i <- 0
for(feature in features.to.plot){
  i <- i + 1
  fb.plot <- FeaturePlot(mm.mog.integ.data, feature , min.cutoff = 0) + 
    ggtitle(paste0('PBMC (GSE155006) - ', feature)) + 
    scale_colour_gradient(breaks = c(0,breaks[i]), low = 'grey85', high = 'blue') + 
    scale_x_continuous(breaks=c(-15, 0, 10)) +
    scale_y_continuous(breaks=c(-15, 0, 5)) +
    theme(plot.title = element_text(hjust = 0.5))
  print(fb.plot) 
  
  tiff(paste0('Figure_5/GSE155006_FeaturePlot_', feature, '.tiff'), width = 7, height = 6, units = 'in', res = 200)
  print(fb.plot)
  dev.off()
  
  fb.plot.blank <- FeaturePlot(mm.mog.integ.data, feature, min.cutoff = 0) + 
    ggtitle('') + xlab('') + ylab('') +
    scale_colour_gradient(breaks = c(0,breaks[i]), low = 'grey85', high = 'blue') +
    scale_x_continuous(breaks=c(-15, 0, 10)) +
    scale_y_continuous(breaks=c(-15, 0, 5)) +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x=element_blank(),
          axis.text.y=element_blank(),
          legend.title=element_blank(),
          legend.text=element_blank())
  print(fb.plot.blank) 
  
  tiff(paste0('Figure_5/GSE155006_FeaturePlot_', feature, '_blank.tiff'), width = 6.5, height = 6, units = 'in', res = 200)
  print(fb.plot.blank)
  dev.off()
}
```